The Calibration Reference Data System

### 2016-03-15 docs/paper/response.txt

A section at the end has been added that briefly describes what Spitzer, 
Chandra and Gemini use and notes some of the restrictions entailed in their
approaches (Chandra's is most similar in approach but simpler).

As far as the implications regarding databases go, we have not expanded on that.
This is mainly because some issues are arguably self evident (distributing
software that requires databases certainly does impose a significantly larger 
burden on those installing and maintaining such a system; we doubt that is 
controversial).

We also think it is nearly self evident that SQL does not always represent 
rules in a clear or easy to understand way (and many astronomers, the ones 
that ultimately must provide those rules, are not experts in SQL, particularly
complex SQL). We aren't saying it is impossible to use only SQL, just not 
ideal. Indeed, Gemini's experience found it to be too constricting (as we did).
Chandra doesn't use it at all, and it appears Spitzer's needs are not as complex.
This is probably aided by the fact that they generally process the data in 
large runs, i.e,. their updates are relatively infrequent as we understand 
it.  Besides, they sometimes have special processing after the query (e.g., 
interpolation between two reference files), so it isn't a pure database
solution either. Accomodating new needs requires new software changes.
To be fair, so does ours if a new kind of rule is required, but that
is much less likely, and much more likely to be in adding a new module
rather than changing existing code.

As to the focus on STScI needs. It would be nice if we had the resources 
to develop a system with a very broad reach and goals initially (well, perhaps,
many very ambitious projects die for being too ambitious!). But with the
reality of funding oversight it's very difficult to do that. NASA gets 
quite nervous when they see work being done that isn't specifically done
for their mission. We do what we can to keep the project more broadly 
useful within these constraints. That's even true when writing papers
(i.e., can we justify spending much time doing a thorough survey of all
systems out there?).

@@

Section 2
We have reworded to be more explicit that all possible combinations of parameters
must be represented as distinct rows.

Section 3 point 2: "Could an 'is valid' boolean attribute in the database
help out?"
Not really. The file itself may be perfectly good, it is just associated with
the wrong value of parameters. Possbily if coupled with a resubmission with the
right values, but note that may mean many entries have to be so marked, and 
replicated. But as implemented, it didn't have that option. If it had been 
designed from the beginning to handle that, yes, it could have dealt with bad
submissions that way.

Section 5: first sentence.
Fixed.
Section 5 submission typo
Fixed
Section 5 point 3
This text has been made clearer by replacing pseudo-mode with more
specific text.

### 2016-03-15 docs/paper/crds.tex
updated all text except for pseudo-mode clarification
git-svn-id: https://aeon.stsci.edu/ssb/svn/crds_server/trunk@2249 37088bc5-bd3f-4c5b-b9de-119bb1e8f7e3

\begin{enumerate}
\item Supporting the bestref functionality.
\item Supporting the submission of new reference files and all the issues that
go with that.
\item Supporting format, semantic, and parameter constraint checking for both
references and rules files.  Having the capability to detect typos, duplicate 
lines, and grammatical errors in rules files. 
Having the capability to analyze basic properties
such as duplicate or deleted rows,  by pseudo-mode, for supported tables files.
\item Providing utilities for making CRDS useful for operations, instrument
scientists that must generate and monitor reference files, and end users.
\end{enumerate}

@@

\subsubsection{Match}
This selector finds which tuple of parameters matches the supplied tuple
of values for those parameters. This is a very flexible selector, and 
many options are provided for its use. (If only one parameteter is being,
matched, the tuple can be replaced by a simple string.)

@@

\begin{description}
\item[Lists of Allowable Alternatives:] Allowable values for parameteters
may be supplied as an \textquotesingle OR\textquotesingle
{\tt }list using {\tt |} as a separator. For example:
\texttt {({\ttfamily"}filter1|filter2{\ttfamily"},...)} will allow either value to match the first
parameter.
\item[Wild Cards:] Using an {\tt *} in a string applies familiar wildcard 
matching rules to the supplied string. For example 
\texttt {({\ttfamily"}F*122{\ttfamily"})} will
match any value starting with {\tt F} and ending with {\tt 122}
\item[Regular Expressions:] This allows matching true regular expressions,
allowing for more complex matching rules than simple wildcards permit.
\item[Literal Expressions:] A literal expression is bracketed with curly
braces {\tt \{ \}} and allows inclusion of special characters without 
special interpretation by CRDS, such as {\tt | and \*}.
\item[Relational Expressions:] Relational expressions are bracketed by {\tt \#}.
These do numerical comparisons on the supplied parameter value to determine
a match. Supported comparison operators are: 
\texttt {\textgreater, \textgreater =, \textless, \textless =, == and or}. 
For example \texttt {({\ttfamily"}\# >=1 and <37",)} 
will match any number equal to or 
greater than 1 and less than 37.
\item[Between:] This is simply a more convenient way of expressing a range
than using relational expressions. E.g., equivalent to the previous example
is: \texttt{{\ttfamily"}between 1 37{\ttfamily"}}. It is inclusive of the first number.
\item[N/A:] This is equivalent to using a wildcard for the value, but with 
the difference of how weighting is used (discussed below). 
\item[NOT Expressions:] A negation of a match. 
\texttt {{\ttfamily"}not HRC{\ttfamily"}} 
is a match if the value does not match {\tt HRC}.
\item[Substitutions:] Substitutions can be defined in the header. Essentially
they result in the replacement of a string in a rules file with a different
string. This is useful in the case where the string is complex and must be 
entered in many different rules (e.g., it is a combination of many different 
allowable values), or where it is more mnemonic to those that must maintain
the rules than the actual header values are.
\end{description}

@@

Gemini uses a hybrid approach, using a database to store much of the 
information, but where much of the selection of calibration reference 
files is performed by running Python code to generate the appropriate 
SQL query to apply to the database (Paul Hirst, private communication). 
The use of pure SQL queries did not capture all the needs of the 
complex selection rules required; their hybrid approach allows arbitrarily
complex rules to be constructed. It does not permit easy distribution of the 
system though it does provide a web-based service to provide calibration 
reference file selections to external users. The results rank multiple 
possible files that may be used as to their applicability. It does try 
to allow for multiple kinds of processing that may require different 
versions of calibration files depending on the kind of processing used 
(e.g., whether you have done an overscan subtraction will affect which 
bias file you are given). As such, it isn’t easy to support multiple 
versions of the rules (that requires multiple servers), nor can the 
system be easily distributed or customized without a lot of expertise 
(but it can be used remotely through the web, and there is consideration 
being given to mechanisms to allow local customization to the web results). 
Reversion is possible by using version control to reinstate the 
appropriate Python modules, but has never been used in practice.

>>>>

\begin{enumerate}
\item Supporting the bestref functionality.
\item Supporting the submission of new reference files and all the issues that
go with that.
\item Supporting format, semantic, and parameter constraint checking for both
references and rules files.  Having the capability to detect typos, duplicate 
lines, and grammatical errors in rules files. 
Having the capability to analyze basic properties
such as duplicate or deleted rows, as defined by critical row lookup parameters 
for supported tables files.
\item Providing utilities for making CRDS useful for operations, instrument
scientists that must generate and monitor reference files, and end users.
\end{enumerate}

@@

\subsubsection{Match}
This selector finds which tuple of parameters matches the supplied tuple
of values for those parameters. This is a very flexible selector, and 
many options are provided for its use. (If only one parameter is being,
matched, the tuple can be replaced by a simple string.)

@@

\begin{description}
\item[Lists of Allowable Alternatives:] Allowable values for parameters
may be supplied as an \textquotesingle OR\textquotesingle
{\tt }list using {\tt |} as a separator. For example:
\texttt {({\ttfamily"}filter1|filter2{\ttfamily"},...)} will allow either value to match the first
parameter.
\item[Wild Cards:] Using an {\tt *} in a string applies familiar wildcard 
matching rules to the supplied string. For example 
\texttt {({\ttfamily"}F*122{\ttfamily"})} will
match any value starting with {\tt F} and ending with {\tt 122}
\item[Regular Expressions:] This allows matching true regular expressions,
allowing for more complex matching rules than simple wildcards permit.
\item[Literal Expressions:] A literal expression is bracketed with curly
braces {\tt \{ \}} and allows inclusion of special characters without 
special interpretation by CRDS, such as {\tt | and \*}.
\item[Relational Expressions:] Relational expressions are bracketed by {\tt \#}.
These do numerical comparisons on the supplied parameter value to determine
a match. Supported comparison operators are: 
\texttt {\textgreater, \textgreater =, \textless, \textless =, == and or}. 
For example \texttt {({\ttfamily"}\# >=1 and <37",)} 
will match any number equal to or 
greater than 1 and less than 37.
\item[Between:] This is simply a more convenient way of expressing a range
than using relational expressions. E.g., equivalent to the previous example
is: \texttt{{\ttfamily"}between 1 37{\ttfamily"}}. It is inclusive of the first number.
\item[N/A:] This is equivalent to using a wildcard for the value, but with 
the difference of how weighting is used (discussed below). 
\item[NOT Expressions:] A negation of a match. 
\texttt {{\ttfamily"}not HRC{\ttfamily"}} 
is a match if the value does not match {\tt HRC}.
\item[Substitutions:] Substitutions can be defined in the header. Essentially
they result in the replacement of a string in a rules file with a different
string. This is useful in the case where the string is complex and must be 
entered in many different rules (e.g., it is a combination of many different 
allowable values), or where it is more mnemonic to those that must maintain
the rules than the actual header values are.
\end{description}

@@

Gemini uses a hybrid approach, using a database to store much of the 
information, but where much of the selection of calibration reference 
files is performed by running Python code to generate the appropriate 
SQL query to apply to the database (Paul Hirst, private communication). 
The use of pure SQL queries did not capture all the needs of the 
complex selection rules required; their hybrid approach allows arbitrarily
complex rules to be constructed. It does not permit easy distribution of the 
system though it does provide a web-based service to provide calibration 
reference file selections to external users. The results rank multiple 
possible files that may be used as to their applicability. It does try 
to allow for multiple kinds of processing that may require different 
versions of calibration files depending on the kind of processing used 
(e.g., whether you have done an overscan subtraction will affect which 
bias file you are given).

As such, it isn’t easy to support multiple 
versions of the rules (that requires multiple servers), nor can the 
system be easily distributed or customized without a lot of expertise 
(but it can be used remotely through the web, and there is consideration 
being given to mechanisms to allow local customization to the web results). 
Reversion may be possible by using version control to reinstate the 
appropriate Python modules, but has never been used in practice; 
for the results to be replicated accurately requires that none of the
information in the database for files present at that time of the previous
results has been changed since that time.
